Data Specialization for a Staged Lambda Calculus

### 2014-11-05 paper/splitting.tex
comments entered

@@

Splitting stage \bbtwo\ terms, as implemented by the judgment $e \splittwosym
[p,l.r]$ (\cref{fig:termSplitTwo}), is guided by the structure of the term $e$.
For example, the stage \bbtwo\ term
\begin{lstlisting}
2`hold{`1` 1+2 `2`} < (hold{`1` 3+4 `2`} - 5)`
\end{lstlisting}
has two subterms, the latter of which itself has two subterms. Two are $\pause$s
containing stage \bbone\ computations, while the third has no stage \bbone\
content. The precomputation \emph{lifts} these computations in the same tree
structure: \verb|(1+2,(3+4,()))|.%
\footnote{In later examples, we will optimize out unnecessary \texttt{()}s.}
The resumer is a function which computes the stage \bbtwo\ value of the original
term by mapping the results of the precomputation in place of the $\pause$s:
\verb|(L1,(L2,_)) => L1 < (L2 - 5)|.

In general, constants and variables have no stage \bbone\ content, so their
precomputations are \texttt{()}, and their resumers return the original term's
value. Non-staging constructs with $n$ subterms have precomputations which are
$n$-tuples of their subterms' precomputations, and resumers which pass each
component to the resumer of the respective subterm. (This behavior at
\texttt{case} and functions causes precomputations to be lifted out from
underneath stage \bbtwo\ binders.) The $\prev$ and $\pause$ rules depends on
splitting stage \bbone\ terms, which is covered later.

@@

We formalize this notion via two \emph{masking} functions (\ref{fig:valSplit}),
$\masko{\cdot}$ and $\maskt{\cdot}$, which expose only the relevant structure of
a partial value at each stage, essentially by replacing all stage \bbone\
(resp., \bbtwo) components of a partial value by \texttt{()}. The first result
of splitting must then compute the \bbone-mask of the partial value, and the
second result the \bbtwo-mask.

@@

>>>>

\ur{Why is there a free variable ``Gamma'' in the figure for splitting
  stage 2?}

@@

\ur{I would not say that ``The correctness of splitting requires'',
  because we don't know what splitting is.  Instead, it is better to
  state this actively, e.g., say that we wish to split the code such
  that the condition e = let l = p in r holds.  correctness will have
  to go beyond this, by saying for example that all stage 1 parts are
  in p etc.  Note that this correctness allows a trivial split
   that shoves everything to r, e.g., p = () and r is ``e''
  (might need to flatten).}

Splitting stage \bbtwo\ terms, as implemented by the judgment $e \splittwosym
[p,l.r]$ (\cref{fig:termSplitTwo}), is guided by the structure of the term $e$.

\ur{At this point, I would have expected a brief discussion of the
  easy cases, like the paragraph that starts with ``In general'' below
  does. You can move that paragraph here and edit to play this role.
  You can then move onto the example below after stating that more
  interesting things happen with prev and hold.}

For example, the stage \bbtwo\ term
\begin{lstlisting}
2`hold{`1` 1+2 `2`} < (hold{`1` 3+4 `2`} - 5)`
\end{lstlisting}
has two subterms, the latter of which itself has two subterms. Two are $\pause$s
containing stage \bbone\ computations, while the third has no stage \bbone\
content.

\ur{The discussion of the subterms is not helpful. what is the point?}

The precomputation \emph{lifts} these computations in the same tree
structure: \verb|(1+2,(3+4,()))|.%
\ur{Not clear what a ``tree structure'' is. }
\footnote{In later examples, we will optimize out unnecessary \texttt{()}s.}
The resumer is a function which computes the stage \bbtwo\ value of the original
term by mapping the results of the precomputation in place of the $\pause$s:
\verb|(L1,(L2,_)) => L1 < (L2 - 5)|.

\ur{The above example should refer to Figure 11 (not Figure 10).  But
  this is not mentioned.}

\ur{This paragraph should come first, followed by a paragraph that
  talks about Figure 11.}
In general, constants and variables have no stage \bbone\ content, so their
precomputations are \texttt{()}, and their resumers return the original term's
value. Non-staging constructs with $n$ subterms have precomputations which are
$n$-tuples of their subterms' precomputations, and resumers which pass each
component to the resumer of the respective subterm. (This behavior at
\texttt{case} and functions causes precomputations to be lifted out from
underneath stage \bbtwo\ binders.) The $\prev$ and $\pause$ rules depends on
splitting stage \bbone\ terms, which is covered later.

@@

\ur{So what goes wrong?  I have no idea what you are talking about at
  this point.  The thought seems unfinished.}

\ur{The following paragraph makes no sense to me, especially because
  it does not match Figure 11.  AS far as I can see Figure 11 does not
talk about masking but value splitting.}
We formalize this notion via two \emph{masking} functions (\ref{fig:valSplit}),
$\masko{\cdot}$ and $\maskt{\cdot}$, which expose only the relevant structure of
a partial value at each stage, essentially by replacing all stage \bbone\
(resp., \bbtwo) components of a partial value by \texttt{()}. The first result
of splitting must then compute the \bbone-mask of the partial value, and the
second result the \bbtwo-mask.

\ur{The following paragraph makes no sense to me, especially because
  it does not match Figure 11.}

@@

\ur{This paragraph is unintellegible to me.  What is the point?}
